---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

## Set Up 

Load required libraries.

```{r}
library(tidyverse)
library(ggpubr)
```

Load required .csv file. This file is organized in a way such that all effect sizes are listed in the file once with a label in the significance column "All", and all significant effect sizes are listed a second time with the label in the significance column "Yes". This is done so that later we can facet the figure by 1) all results regardless of significance and 2) only significant results. 


```{r}
effectsize_final <- read.csv("Final Spreadsheet/effectsizesforfig_final.csv", stringsAsFactors = TRUE, na.strings=c(""))

effectsize_final$All.Logs<-as.numeric(as.character(effectsize_final$All.Logs))

str(effectsize_final)

```


## Compositional effect sizes only

Filter the initial dataset by noncomp results only. 

```{r}

effectsize_final_comp <- effectsize_final %>% 
  dplyr::filter(Type == "Comp")
  
```

Generate a separate dataset that is significant results only for the dot plot.

```{r}
effectsize_c_orders_sig <-
  effectsize_final_comp %>% 
  dplyr::filter(Sig == "Yes") %>% 
  dplyr::filter(Order == "Lactobacillales" | Order == "Orbales" | Order == "Bifidobacteriales" | Order == "Neisseriales") %>%
  dplyr::filter(Core == "Y") %>% 
  dplyr::filter(All.Logs>-50 & All.Logs<50) %>% ##get rid of extreme outliers
  dplyr::filter(!is.na(All.Logs))

fig_c_effectsizeorder <- effectsize_final_comp %>% 
  dplyr::filter(!is.na(All.Logs)) %>%  #get rid of NAs
  dplyr::filter(Order == "Lactobacillales" | Order == "Orbales" | Order == "Bifidobacteriales" | Order == "Neisseriales") %>% 
  dplyr::filter(Core == "Y") %>% 
  dplyr::filter(All.Logs>-50 & All.Logs<50) %>% ##get rid of extreme outliers
  ggplot(aes(x=Sig, y = All.Logs)) +
  geom_rect(xmin = 0, xmax = 1.5, ymin = -Inf, ymax = Inf, fill = "#DCDCDC")+
  geom_boxplot(position=position_dodge(width = 2)) +
  geom_dotplot(data = effectsize_c_orders_sig, stackgroups = TRUE, method = "histodot",
               binaxis = 'y', stackdir = 'center', dotsize = 1, 
               position = position_dodge(width = 0), 
               aes(fill= Pesticide),  alpha=0.7)+
  scale_fill_manual(name = "Pesticide",values=c("Chlorothalonil" = "#8F2D56",    
                                                "Coumaphos" = "#EE2F75",
                                                "Fipronil" = "#ED6D5A",
                                                "Glyphosate" = "#FFBC42",
                                                "Imidacloprid" = "#00BD9D",
                                                "Nitenpyram" = "#047815",
                                                "Pristine" = "#063000",
                                                "Roundup" = "#00A9FE",
                                                "Spinosad" = "#0050D9",
                                                "Thiamethoxam" = "#7C18AF"))+
  geom_hline(aes(yintercept = 0), colour = "#696969", linetype = "longdash")+
  geom_hline(aes(yintercept = 1), colour = "#696969", linetype = "dashed")+
  geom_hline(aes(yintercept = -1), colour = "#696969", linetype = "dashed")+
  theme_classic() +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    legend.title = element_blank(),
    strip.placement = "outside", 
    strip.text.x = element_text( #text in facet labels
      size = 12, color = "black", face = "bold"),
    strip.background = element_rect(#background of facet labels
      color="white", fill="#ffffff", size=1.5),
    panel.spacing.y = unit(2, "lines"),
    legend.position = "bottom",
    axis.title.y = element_blank())+
  theme(text = element_text(size = 14), axis.text.y = element_text(color="black", size = 10))+
  scale_y_continuous(name ="Log2 Fold Change \n(Treatment/Control)")+
  scale_x_discrete(labels=c("All" = "All Results", "Yes" = "Significant \nResults"), expand = c(0.5, 0))+
  facet_wrap(~ Order, scales = 'free', ncol = 2, strip.position = "top")

```

Then make another panel for the two Lactobacillales phylotypes.

```{r}
effectsize_c_lacto_sig <- 
  effectsize_final_comp %>% 
  dplyr::filter(Sig == "Yes") %>% 
  dplyr::filter(Species == "Lactobacillus Firm 4" | Species == "Lactobacillus Firm 5") %>%
  dplyr::filter(Core == "Y") %>% 
  dplyr::filter(All.Logs>-50 & All.Logs<50) %>% ##get rid of extreme outliers
  dplyr::filter(!is.na(All.Logs))

fig_c_effectsizelacto <- effectsize_final_comp %>% 
  dplyr::filter(!is.na(All.Logs)) %>% #get rid of NAs
  dplyr::filter(Species == "Lactobacillus Firm 4" | Species == "Lactobacillus Firm 5") %>%
  dplyr::filter(Core == "Y") %>% 
  dplyr::filter(All.Logs>-50 & All.Logs<50) %>% ##get rid of extreme outliers
  ggplot(aes(x=Sig, y = All.Logs)) +
  geom_rect(xmin = 0, xmax = 1.5, ymin = -Inf, ymax = Inf, fill = "#DCDCDC")+
  geom_boxplot(position=position_dodge(width = 2)) + 
  geom_dotplot(data = effectsize_c_lacto_sig, stackgroups = TRUE, method = "histodot",
               binaxis = 'y', stackdir = 'center', dotsize = 1.2, 
               position = position_dodge(width = 0), 
               aes(fill= Pesticide),  alpha=0.7)+
  scale_fill_manual(name = "Pesticide",values=c("Chlorothalonil" = "#8F2D56",    
                                                "Coumaphos" = "#EE2F75",
                                                "Fipronil" = "#ED6D5A",
                                                "Glyphosate" = "#FFBC42",
                                                "Imidacloprid" = "#00BD9D",
                                                "Nitenpyram" = "#047815",
                                                "Pristine" = "#063000",
                                                "Roundup" = "#00A9FE",
                                                "Spinosad" = "#0050D9",
                                                "Thiamethoxam" = "#7C18AF"))+
  geom_hline(aes(yintercept = 0), colour = "#696969", linetype = "longdash")+
  geom_hline(aes(yintercept = 1), colour = "#696969", linetype = "dashed")+
  geom_hline(aes(yintercept = -1), colour = "#696969", linetype = "dashed")+
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.title = element_blank(),
        strip.placement = "outside", 
        strip.text.x = element_text( #text in facet labels
          size = 12, color = "black", face = "bold"),
        strip.background = element_rect(#background of facet labels
          color="white", fill="#ffffff", size=1.5),
        panel.spacing.y = unit(2, "lines"),
        legend.position = "none",
        axis.title.y = element_blank())+
  theme(text = element_text(size = 14), axis.text.x = element_text(vjust=0.3, size = 12, color="black"), axis.text.y = element_text(color="black", size = 10))+
  scale_y_continuous(name ="Log2 Fold Change \n(Treatment/Control)")+
  scale_x_discrete(labels=c("All" = "All Results", "Yes" = "Significant \nResults"), expand = c(0.5, 0))+
  facet_wrap(~ Species, scales = 'free', ncol = 2, strip.position = "top")
```

## Non-compositional effect sizes only

Filter the initial dataset by noncomp results only. 

```{r}

effectsize_final_noncomp <- effectsize_final %>% 
  dplyr::filter(Type == "Noncomp")

```

Now use the same code as before, just with the noncomp dataset. 

```{r}
effectsize_nc_orders_sig <-
  effectsize_final_noncomp %>% 
  dplyr::filter(Sig == "Yes") %>% 
  dplyr::filter(Order == "Lactobacillales" | Order == "Orbales" | Order == "Bifidobacteriales" | Order == "Neisseriales") %>%
   dplyr::filter(All.Logs>-50 & All.Logs<50) %>% ##get rid of extreme outliers
  dplyr::filter(Core == "Y") %>% 
  dplyr::filter(!is.na(All.Logs))

fig_nc_effectsizeorder <- effectsize_final_noncomp %>% 
  dplyr::filter(!is.na(All.Logs)) %>%  #get rid of NAs
  dplyr::filter(Order == "Lactobacillales" | Order == "Orbales" | Order == "Bifidobacteriales" | Order == "Neisseriales") %>% 
   dplyr::filter(All.Logs>-50 & All.Logs<50) %>% ##get rid of extreme outliers
  dplyr::filter(Core == "Y") %>% 
  ggplot(aes(x=Sig, y = All.Logs)) +
  geom_rect(xmin = 0, xmax = 1.5, ymin = -Inf, ymax = Inf, fill = "#DCDCDC")+
  geom_boxplot(position=position_dodge(width = 2)) +
  geom_dotplot(data = effectsize_nc_orders_sig, stackgroups = TRUE, method = "histodot",
               binaxis = 'y', stackdir = 'center', dotsize = 1, 
               position = position_dodge(width = 0), 
               aes(fill= Pesticide),  alpha=0.7)+
  scale_fill_manual(name = "Pesticide",values=c("Chlorothalonil" = "#8F2D56",    
                                                "Coumaphos" = "#EE2F75",
                                                "Fipronil" = "#ED6D5A",
                                                "Glyphosate" = "#FFBC42",
                                                "Imidacloprid" = "#00BD9D",
                                                "Nitenpyram" = "#047815",
                                                "Pristine" = "#063000",
                                                "Roundup" = "#00A9FE",
                                                "Spinosad" = "#0050D9",
                                                "Thiamethoxam" = "#7C18AF"))+
  geom_hline(aes(yintercept = 0), colour = "#696969", linetype = "longdash")+
  geom_hline(aes(yintercept = 1), colour = "#696969", linetype = "dashed")+
  geom_hline(aes(yintercept = -1), colour = "#696969", linetype = "dashed")+
  theme_classic() +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    legend.title = element_blank(),
    strip.placement = "outside", 
    strip.text.x = element_text( #text in facet labels
      size = 12, color = "black", face = "bold"),
    strip.background = element_rect(#background of facet labels
      color="white", fill="#ffffff", size=1.5),
    panel.spacing.y = unit(2, "lines"),
    legend.position = "bottom",
    axis.title.y = element_blank())+
  theme(text = element_text(size = 14),  axis.text.y = element_text(color="black", size = 10))+
  scale_y_continuous(name ="Log2 Fold Change \n(Treatment/Control)")+
  scale_x_discrete(labels=c("All" = "All Results", "Yes" = "Significant \nResults"), expand = c(0.5, 0))+
  facet_wrap(~ Order, scales = 'free', ncol = 2, strip.position = "top")

```

Then the Lactobacillus phylotypes specific panel.

```{r}
effectsize_nc_lacto_sig <- 
  effectsize_final_noncomp %>% 
  dplyr::filter(Sig == "Yes") %>% 
  dplyr::filter(Species == "Lactobacillus Firm 4" | Species == "Lactobacillus Firm 5") %>%
  dplyr::filter(Core == "Y") %>% 
   dplyr::filter(All.Logs>-50 & All.Logs<50) %>% ##get rid of extreme outliers
  dplyr::filter(!is.na(All.Logs))

fig_nc_effectsizelacto <- effectsize_final_noncomp %>% 
  dplyr::filter(!is.na(All.Logs)) %>% #get rid of NAs
  dplyr::filter(Species == "Lactobacillus Firm 4" | Species == "Lactobacillus Firm 5") %>%
  dplyr::filter(Core == "Y") %>% 
   dplyr::filter(All.Logs>-50 & All.Logs<50) %>% ##get rid of extreme outliers
  ggplot(aes(x=Sig, y = All.Logs)) +
  geom_rect(xmin = 0, xmax = 1.5, ymin = -Inf, ymax = Inf, fill = "#DCDCDC")+
  geom_boxplot(position=position_dodge(width = 2)) + 
  geom_dotplot(data = effectsize_nc_lacto_sig, stackgroups = TRUE, method = "histodot",
               binaxis = 'y', stackdir = 'center', dotsize = 1.2, 
               position = position_dodge(width = 0), 
               aes(fill= Pesticide),  alpha=0.7)+
  scale_fill_manual(name = "Pesticide",values=c("Chlorothalonil" = "#8F2D56",    
                                                "Coumaphos" = "#EE2F75",
                                                "Fipronil" = "#ED6D5A",
                                                "Glyphosate" = "#FFBC42",
                                                "Imidacloprid" = "#00BD9D",
                                                "Nitenpyram" = "#047815",
                                                "Pristine" = "#063000",
                                                "Roundup" = "#00A9FE",
                                                "Spinosad" = "#0050D9",
                                                "Thiamethoxam" = "#7C18AF"))+
  geom_hline(aes(yintercept = 0), colour = "#696969", linetype = "longdash")+
  geom_hline(aes(yintercept = 1), colour = "#696969", linetype = "dashed")+
  geom_hline(aes(yintercept = -1), colour = "#696969", linetype = "dashed")+
  theme_classic() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.title = element_blank(),
        strip.placement = "outside", 
        strip.text.x = element_text( #text in facet labels
          size = 12, color = "black", face = "bold"),
        strip.background = element_rect(#background of facet labels
          color="white", fill="#ffffff", size=1.5),
        panel.spacing.y = unit(2, "lines"),
        legend.position = "none",
        axis.title.y = element_blank())+
  theme(text = element_text(size = 14), axis.text.x = element_text(vjust=0.3, size = 12, color="black"), axis.text.y = element_text(color="black", size = 10))+
  scale_y_continuous(name ="Log2 Fold Change \n(Treatment/Control)")+
  scale_x_discrete(labels=c("All" = "All Results", "Yes" = "Significant \nResults"), expand = c(0.5, 0))+
  facet_wrap(~ Species, scales = 'free', ncol = 2, strip.position = "top")
```


Now make a four panel figures from all the plots. 

```{r}
effectsizesfig_fourpanel <-
  ggarrange(fig_c_effectsizeorder, 
            fig_nc_effectsizeorder,
            fig_c_effectsizelacto,
            fig_nc_effectsizelacto,
            heights = c(2, 1), nrow=2, ncol = 2, common.legend = TRUE, legend = "bottom"
           )

effectsizesfig_fourpanel
```

Annotate the figure so that it has a common y-axis. 

```{r}
annotate_figure(effectsizesfig_fourpanel,
                left = ggpubr::text_grob(expression(""*Log[2]~Fold~Change ~(Treatment/Control)*""), 
                  size = 16, color = "black", rot = 90))

```

